### 2018-04-06 paper.Rmd
typos and other edits

author_note: |
  Thanks to Brian MacWhinney for advice and guidance, and to Melissa Kline for her work on ClanToR, which formed a starting point for our work. This work supported by a Jacobs Advanced Research Fellowship to MCF.

@@

Established in 1984 to address this issue, the Child Language Data Exchange System (CHILDES) aims to make transcripts and recording of children's language available for researchers via a free resource [@macwhinney2000]. CHILDES archives tens of thousands of transcripts of children’s speech, making it an critical resource for characterizing children’s early productive language use as well as their language environment. As the first major effort to consolidate and share transcripts of child language, CHILDES has been a pioneer in the move to curate and disseminate large-scale behavioral datasets publicly.

@@

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal. CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- can be tricky to build into a processing pipeline written in a modern analytic platform such as python or R. Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analysis typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998;@yang2013;@meylan2017].

This practice is problematic for a number of reasons. Effort wasted in rewriting the same parser again and again. In addition, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard. Further, these scripts are rarely shared, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017]. This lack of computational reproducibility is widely recognized as problematic because it does not allow researchers to verify or build on published researcher [@donoho2010;@stodden2016]. Finally, the CHILDES corpus itself is a moving target;  computational work using the entire corpus at one time point may include a different set of data than subsequent work (due to new contributions or revisions).

In the current manuscript, we describe a system for extending the functionality of CHILDES to address these issues. Our systemn, childes-db, is a database-formatted mirror of CHILDES that allows programmatic access through an application programming interface (API). This infrastructure allows the creation of web applications for browsing and visualizing the data easily, facilitating classroom use of the dataset. Further, it can be accessed programmatically by advanced researchers, obviating the need to write one-off parsers of the CHAT format. The database is versioned for access to previous images, allowing computational reproducibility of particular analyses.

@@

Unlike the *parsing* approach in CLAN which entails the sequential processing of strings, childes-db treats CHILDES as a set of linked tables, with records corresponding to intuitive abstractions such as words, utterances, and transcripts. Users of data analysis languages like R or Julia, libraries like Pandas, or those familiar with Structured Query Language (SQL) will be familiar with operations on tables such as filtering (subsetting), sorting, aggregation (grouping), and joins (merges). We expect most users to interact with the database through the `childesr` API for the `R` language described below, we begin with a top-level description and motivation for the design of the database schema, details of the database’s current technical implementation, and our versioning scheme.

At its core, childes-db is a database consisting of a set of linked tabular data stores where rows correspond to records. The smallest unit of abstraction tracked by the database is a token, treated here as the citation orthographic form of a word. Using the standardized written form of the word facilitates the computation of lexical frequency statistics, for comparison or aggregation across children or time periods. Deviations from the citation form -- reflecting the deviations from citation forms that are particularly common in the course of language development -- are kept as a separate (possibly null) field associated with each token.

Other tables in the database comprise a hierarchical system of collections built on top of tokens -- *utterance*, *transcription*, *corpus*, and *collection* -- which stores attributes appropriate for each level of description. Every entity includes attributes that link it to all higher-order collections, e.g., an utterance lists the transcript, corpus, and collection to which it belongs.  An *utterance* contains one or more words and includes fields such as the utterance type (declarative, question, etc.), total number of tokens, and the total number one morphemes.

A *transcription* consists of one or more utterances and includes the date collected, the name of the target child and age in days (if defined), and the filename from CHILDES. A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpus. Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES. Because every entity can be linked to a top-level collection, childes-db places content from all languages in the same table for a given level of abstraction.

@@

To reflect changes in the CHILDES database as new corpora are added or previous annotations revised, we produce a new versioned parse of the complete database every [[n]] months. By default, users connect to the most recent version. To support reproduction of results with previous versions of the database, we provide hosted versions of the last [[m]] versions so that researchers can refer to specific releases. Compressed .sql files are available for download for [[r]] years.

@@

One common use of the CHILDES corpora is estimating the frequency with which children hear different words. These frequency estimates are used both in the development of theory [e.g. frequent words are learned earlier; @goodman2008], and in the construction of age-appropriate experimental stimuli. One benefit of the `childesdb` interface is that it allows for easy analysis of how the frequencies of words change over development. Many of our theories in which children learn the structure of language from its statistical properties implicitly assume that these statistics are *stationary*, i.e. unchanging over development [e.g., @saffran1996]. However a number of recent analysis show that the frequencies with which infants encounter both linguistic and visual properties of their environment may change dramatically over development [@fausey2016], and these changing distributions may produce similarly dramatic changes in the ease or difficulty with which these regularities can be learned [@elman1993].

To demonstrate how one might discover such a non-stationary, we take as a case study the frequency with which children hear the color words of English (e.g. "blue", "green"). Color words tend to be learned relatively late by children, potentially in part due to the abstractness of the meanings to which they refer [see @wagner2013]. However, within the set of color words, the frequency with which these words are heard predicts a significant fraction of the variance in their order of acquisition [@yurovsky2015]. But are these frequencies stationary--e.g. do children hear "blue" as often at 12 months as they do at 24 months? We answer this question in two ways--first using the childesb shiny apps, and then using the `childesr` package.

@@

To investigate whether the frequency of color words is stationary over development, a user can navigate to the Frequency app, and enter a set of color words into the `Word` selector separated by a comma: here "blue, red, green." Because the question of interest is about the frequency of words in the input (rather than produced by children), the `Speaker` field can be set to reflect this. In this example we select "Mother" because, across corpora in CHIDLES, the speaker tagged with the Mother role tends to produce most of the words that children hear. Because children learn most of their basic color words by the age of 5, the age range 1-5 years is a reaonable choice for `Ages to include`. The results of these selections are shown in Figure \ref{fig:color-shiny}. We can also create a hyperlink to store these set of choices so that we can share these results with others, or with ourselves in the future, byby clicking on the `Bookmark` button in the bottom left corner.

@@

data <- stats %>%
  filter(!is.na(target_child_sex), language == "eng") %>%
  group_by(target_child_id, target_child_age, target_child_sex, language) %>%
  summarise(ttr = sum(num_types) / sum(num_tokens))

ggplot(data, aes(x = target_child_age, y = ttr, color = target_child_sex)) +
  geom_smooth()
```

>>>>

author_note: |
  Thanks to Brian MacWhinney for advice and guidance, and to Melissa Kline for her work on ClanToR, which formed a starting point for our work. This work is supported by a Jacobs Advanced Research Fellowship to MCF.

@@

Established in 1984 to address this issue, the Child Language Data Exchange System (CHILDES) aims to make transcripts and recording of children's language available for researchers via a free resource [@macwhinney2000]. CHILDES archives tens of thousands of transcripts of children’s speech, making it a critical resource for characterizing children’s early productive language use as well as their language environment. As the first major effort to consolidate and share transcripts of child language, CHILDES has been a pioneer in the move to curate and disseminate large-scale behavioral datasets publicly.

@@

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal. CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- can be tricky to build into a processing pipeline written in a modern analytic platform such as Python or R. Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analysis typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998;@yang2013;@meylan2017].

This practice is problematic for a number of reasons. Effort is wasted in rewriting the same parser again and again. In addition, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard. Further, these scripts are rarely shared, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017]. This lack of computational reproducibility is widely recognized as problematic because it does not allow researchers to verify or build on published research [@donoho2010;@stodden2016]. Finally, the CHILDES corpus itself is a moving target;  computational work using the entire corpus at one time point may include a different set of data than subsequent work (due to new contributions or revisions).

In the current manuscript, we describe a system for extending the functionality of CHILDES to address these issues. Our system, childes-db, is a database-formatted mirror of CHILDES that allows programmatic access through an application programming interface (API). This infrastructure allows the creation of web applications for browsing and visualizing the data easily, facilitating classroom use of the dataset. Further, it can be accessed programmatically by advanced researchers, obviating the need to write one-off parsers of the CHAT format. The database is versioned for access to previous images, allowing computational reproducibility of particular analyses.

@@

Unlike the *parsing* approach in CLAN which entails the sequential processing of strings, childes-db treats CHILDES as a set of linked tables, with records corresponding to intuitive abstractions such as words, utterances, and transcripts. Users of data analysis languages like R or Julia, libraries like Pandas, or those familiar with Structured Query Language (SQL) will be familiar with operations on tables such as filtering (subsetting), sorting, aggregation (grouping), and joins (merges). Although we expect most users to interact with the database through the `childesr` API for the `R` language described below, we begin with a top-level description and motivation for the design of the database schema, details of the database’s current technical implementation, and our versioning scheme.

At its core, childes-db is a database consisting of a set of linked tabular data stores where rows correspond to records. <!-- what did you mean here? don't rows = records--> The smallest unit of abstraction tracked by the database is a token, treated here as the citation orthographic form of a word. Using the standardized written form of the word facilitates the computation of lexical frequency statistics, for comparison or aggregation across children or time periods. Deviations from the citation form -- reflecting the deviations from citation forms that are particularly common in the course of language development -- are kept as a separate (possibly null) field associated with each token.

Other tables in the database comprise a hierarchical system of collections built on top of tokens -- *utterance*, *transcription*, *corpus*, and *collection* -- which stores attributes appropriate for each level of description. Every entity includes attributes that link it to all higher-order collections, e.g., an utterance lists the transcript, corpus, and collection to which it belongs.  An *utterance* contains one or more words and includes fields such as the utterance type (declarative, question, etc.), total number of tokens, and the total number of morphemes (if the morphological structure is available in the original CHAT).

A *transcription* consists of one or more utterances and includes the date collected, the name of the target child, and age in days (if defined), and the filename from CHILDES. A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpus. Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES. Because every entity can be linked to a top-level collection, childes-db places content from all languages in the same table for a given level of abstraction <!-- a little confused by this last sentence -->.

@@

<!-- Potentially put some R code to show easy it is to access previous versions? -->

To reflect changes in the CHILDES database as new corpora are added or previous annotations revised, we produce a new versioned parse of the complete database every 6 months. By default, users connect to the most recent version. To support reproduction of results with previous versions of the database, we provide access to hosted versions through our `childesr` API so that researchers can refer to specific releases. Compressed .sql files are available for download.

@@

One common use of the CHILDES corpora is estimating the frequency with which children hear different words. These frequency estimates are used both in the development of theory [e.g. frequent words are learned earlier; @goodman2008], and in the construction of age-appropriate experimental stimuli. One benefit of the childes-db interface is that it allows for easy analysis of how the frequencies of words change over development. Many of our theories in which children learn the structure of language from its statistical properties implicitly assume that these statistics are *stationary*, i.e. unchanging over development [e.g., @saffran1996]. However a number of recent analyses show that the frequencies with which infants encounter both linguistic and visual properties of their environment may change dramatically over development [@fausey2016], and these changing distributions may produce similarly dramatic changes in the ease or difficulty with which these regularities can be learned [@elman1993].

To demonstrate how one might discover such a non-stationary, we take as a case study the frequency with which children hear the color words of English (e.g. "blue", "green"). Color words tend to be learned relatively late by children, potentially in part due to the abstractness of the meanings to which they refer [see @wagner2013]. However, within the set of color words, the frequency with which these words are heard predicts a significant fraction of the variance in their order of acquisition [@yurovsky2015]. But are these frequencies stationary -- e.g. do children hear "blue" as often at 12 months as they do at 24 months? We answer this question in two ways -- first using the childesb shiny apps, and then using the `childesr` package.

@@

To investigate whether the frequency of color words is stationary over development, a user can navigate to the Frequency app, and enter a set of color words into the `Word` selector separated by a comma: here "blue, red, green." Because the question of interest is about the frequency of words in the input (rather than produced by children), the `Speaker` field can be set to reflect this. In this example we select "Mother" because, across corpora in CHILDES, the speaker tagged with the Mother role tends to produce most of the words that children hear. Because children learn most of their basic color words by the age of 5, the age range of 1-5 years is a reaonable choice for `Ages to include`. The results of these selections are shown in Figure \ref{fig:color-shiny}. We can also create a hyperlink to store these set of choices so that we can share these results with others, or with ourselves in the future, by clicking on the `Bookmark` button in the bottom left corner.

@@

data <- stats %>%
  filter(!is.na(target_child_sex), language == "eng") %>%
  group_by(target_child_id, target_child_age, target_child_sex, language) %>%
  summarise(measure = mean(mtld))

ggplot(data, aes(x = target_child_age, y = measure, color = target_child_sex)) +
  geom_smooth()
```
